---
title: "RTLS Case Study"
author: "Martin Garcia, Michael Catalano, Jeremy Otsap, Christian Nava"
date: "5/19/2020"
output: html_document
---

```{r,echo=FALSE}

library(tidyverse)
library(dplyr)
library(magrittr)
library(lattice)

```

# Data Cleanup

The data must go through several transformations not only for interpretability but for modeling reasons. The format consist of uneven rows of signals from multiple sources received by our scanning device per location and orientation. We split our data into two respective data matrices, one for our scanning device, position, and angle and a separate matrix for signals detected by our device at given angle and position. These matrices are combined to create an dataframe now that we have even number of columns. This creatse a dataframe in long format as we have several signals from the same scanning device at the same location and orientation combination.

```{r, echo=FALSE}
previewoffline = readLines("Data/offline.final.trace.txt")
scantokens = strsplit(previewoffline[4], "[;=,]")[[1]]

```


```{r}
table(scantokens[ - ( 1:10 ) ])

table(scantokens[c(2, 4, 6:8, 10)])

```

We will have to create variables subsets of our dataframe based on different macs of interest. Those that include the highest signal counts as well as those identified as our access points. This will remove the noise from other signals on separate floors. 

```{r}
#Mac Combinations

co= c("00:0f:a3:39:e1:c0", "00:14:bf:b1:97:8a","00:14:bf:3b:c7:c6", "00:14:bf:b1:97:90", "00:14:bf:b1:97:8d","00:14:bf:b1:97:81")
cd = c("00:0f:a3:39:dd:cd", "00:14:bf:b1:97:8a","00:14:bf:3b:c7:c6", "00:14:bf:b1:97:90", "00:14:bf:b1:97:8d","00:14:bf:b1:97:81")
co_cd = c("00:0f:a3:39:e1:c0", "00:0f:a3:39:dd:cd", "00:14:bf:b1:97:8a","00:14:bf:3b:c7:c6", "00:14:bf:b1:97:90", "00:14:bf:b1:97:8d","00:14:bf:b1:97:81")

```

We have several signals for the same combination of access point, angle and location which needs to be summarized by using the mean value. This gives us a single value per location per given combination of access point, angle and location. 

## ProcessLine function
* strips the values separating our data, i.e. ; , =
* combines the scanning device data with each signal detected

```{r,echo=FALSE}

processLine = function(x)
{
tokens = strsplit(x, "[;=,]")[[1]] #strip semicolon, equal and comma of first row
if (length(tokens) == 10)
    return(NULL)
tmp = matrix(tokens[ - (1:10) ], ncol= 4, byrow = TRUE) #pull the last elements except the scanning device (first ten tokens)
cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow(tmp), 6, #column bind the scanning device data with our signal data above
byrow = TRUE), tmp)
}
```

## roundOrientation function
* rounds our angles to the closest 45th angle
```{r}

roundOrientation = function(angles) 
  {
  refs = seq(0, by = 45, length = 9)
  q = sapply(angles, function(o) which.min(abs(o - refs)))
  c(refs[1:8], 0) [q]
  }

```

## readData function
* combines functions above plus other data munging
* names columns
* drops unneeded columns, i.e. z position
* changes time to seconds
```{r}
#Create a function the prep the data
readData = function(filename,  submacs) #filename can be offline or online, macs filter our data to our macs of interest

  {
    txt = readLines(filename)
    lines = txt[ substr(txt, 1, 1) != "#" ]
    tmp = lapply(lines, processLine)
    df = as.data.frame(do.call("rbind", tmp), 
                            stringsAsFactors= FALSE) 
    
    names(df) = c("time", "scanMac", "posX", "posY", "posZ", "orientation", "mac", "signal", "channel", "type")
    
     # keep only signals from access points
    df = df[ df$type == "3", ]
    
    # drop scanMac, posZ, channel, and type - no info in them
    dropVars = c("scanMac", "posZ", "channel", "type")
    df = df[ , !( names(df) %in% dropVars ) ]
    
    # drop more unwanted access points
    df = df[ df$mac %in% submacs, ]
    
    # convert numeric values
    numVars = c("time", "posX", "posY", "orientation", "signal")
    df[ numVars ] = lapply(df[ numVars ], as.numeric)

    # convert time to POSIX
    df$time = df$time/1000
    class(df$time) = c("POSIXt", "POSIXct")
    
    # round orientations to nearest 45
    df$angle = roundOrientation(df$orientation)
    
    #Combine all possible X Y combinations
    df$posXY = paste(df$posX, df$posY, sep = "-")

    return(df)
  }


```

# Model Building

## KNN Test 1: All 7 Mac IDs

```{r}
offline = readData(filename = "Data/offline.final.trace.txt", 
                   submacs  = co_cd)
```

## reshapeSS Function
* function to reshape our data from long to short, creates column of average signal per mac
* note, we apply this function after we take summary level details

```{r}
reshapeSS = function(data, varSignal = "signal", 
                     keepVars = c("posXY", "posX","posY")) {
  byLocation =
    with(data, by(data, list(posXY), 
                  function(x) {
                    ans = x[1, keepVars]
                    avgSS = tapply(x[ , varSignal ], x$mac, mean)
                    y = matrix(avgSS, nrow = 1, ncol = length(unique(offline$mac)),  #reshape for our training data, number of columns dependent on macs, this refers to unique number of macs within the mac column
                               dimnames = list(ans$posXY,
                                               names(avgSS)))
                    cbind(ans, y)
                  }))

  newDataSS = do.call("rbind", byLocation)
  return(newDataSS)
}


```

## Create summary level detail, which takes the mean of every position, angle, and mac combination

```{r}

# Create new combined position variable
offline$posXY = paste(offline$posX, offline$posY, sep="-")
byLocAngleAP = with(offline,
                  by (offline, list (posXY, angle, mac),
                      function(x) x))

# Create a summary of signal strength data - takes combination each unique combination of posxy. angle, mac address to create scalar values of means, medians, etc
# Create a summary of signal strength data
signalSummary = 
  lapply(byLocAngleAP, 
        function(oneLoc) 
        {
          ans = oneLoc[1, ]
          ans$medSignal = median(oneLoc$signal)
          ans$avgSignal = mean(oneLoc$signal)
          ans$num = length(oneLoc$signal)
          ans$sdSignal = sd(oneLoc$signal)
          ans$iqrSignal = IQR(oneLoc$signal)
          ans
        })

# Add signalSummary to the offlineSummary data
offlineSummary = do.call("rbind", signalSummary)
head(offlineSummary)
```

## Test Data

## Clean up our test set, this means rounding orientations, removing ,:= , summarizing data, and changing from long to short
* clean our data again with readData
* online - cleaned version
* onlinesummary - summary level, which translates to the mean of every signal for every mac

```{r}
#Cleans up our data w/ readData
online = readData(filename = "Data/online.final.trace.txt", submacs  = co_cd)

angle_count <- online %>% 
  count(angle)

print(angle_count)

print(str(online))

online$posXY = paste(online$posX, online$posY, sep = "-")

keepVars = c("posXY", "posX","posY", "orientation", "angle")
byLoc = with(online, 
             by(online, list(posXY), 
                function(x) {
                  ans = x[1, keepVars]
                  avgSS = tapply(x$signal, x$mac, mean)
                  y = matrix(avgSS, nrow = 1, ncol = 7, #this is co_cd in submacs
                        dimnames = list(ans$posXY, names(avgSS)))
                  cbind(ans, y)
                }))

onlineSummary = do.call("rbind", byLoc)  
```

## selectTrain function

* function that Subsets our training data to include only specific number of angles and also applies reshapeSS which changes from long to short

```{r}

#Create function that will give us starting angle and range (m)
selectTrain = function(angleNewObs, signals , m){
  # m is the number of angles to keep between 1 and 5
  refs = seq(0, by = 45, length  = 8)
  nearestAngle = roundOrientation(angleNewObs)
  
  if (m %% 2 == 1) 
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
  else {
    m = m + 1
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
    if (sign(angleNewObs - nearestAngle) > -1) 
      angles = angles[ -1 ]
    else 
      angles = angles[ -m ]
  }
  angles = angles + nearestAngle
  angles[angles < 0] = angles[ angles < 0 ] + 360
  angles[angles > 360] = angles[ angles > 360 ] - 360
  angles = sort(angles) 
  
  offlineSubset = signals[ signals$angle %in% angles, ]
  reshapeSS(offlineSubset, varSignal = "avgSignal")
}

```

## KNN function-----------------
* trains on our subset summary level data, 

```{r}
findNN = function(newSignal, trainSubset) {
  diffs = apply(trainSubset[ , 4:10], 1, #looks at the signal data, 10 got cd & co
                function(x) x - newSignal)
  dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) ) #takes differences of signal and squares
  closest = order(dists) #reorders differences
  return(trainSubset[closest, 1:3 ]) #keeps top 3 
}

```

## Prediction Function

```{r}
predXY = function(newSignals, newAngles, trainData, 
                  numAngles = 1, k = 3){
  
  closeXY = list(length = nrow(newSignals))
  
  for (i in 1:nrow(newSignals)) {
    trainSS = selectTrain(newAngles[i], trainData, m = numAngles)
    closeXY[[i]] = 
      findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
  }

  estXY = lapply(closeXY, 
                 function(x) sapply(x[ , 2:3], 
                                    function(x) mean(x[1:k])))
  estXY = do.call("rbind", estXY)
  return(estXY)
}

```

```{r actual knn models}
estXYk1 = predXY(newSignals = onlineSummary[ , 6:12], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 1)

estXYk3 = predXY(newSignals = onlineSummary[ , 6:12], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 3)

estXYk5 = predXY(newSignals = onlineSummary[ , 6:12], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 5)

estXYk7 = predXY(newSignals = onlineSummary[ , 6:12], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 7)
```

## KNN Test 1 Results: Using all 7 MacIDs

```{r pred error}
# Error calculations
calcError = 
  function(estXY, actualXY) 
    sum( rowSums( (estXY - actualXY)^2) )

actualXY = onlineSummary[ , c("posX", "posY")]
sapply(list(estXYk7, estXYk5, estXYk3, estXYk1), calcError, actualXY)
```

## KNN Test 2: Removing cd Mac ID

```{r}
offline = readData(filename = "Data/offline.final.trace.txt", 
                   submacs  = co)
```

## reshapeSS Function
* function to reshape our data from long to short, creates column of average signal per mac
* note, we apply this function after we take summary level details

```{r}
reshapeSS = function(data, varSignal = "signal", 
                     keepVars = c("posXY", "posX","posY")) {
  byLocation =
    with(data, by(data, list(posXY), 
                  function(x) {
                    ans = x[1, keepVars]
                    avgSS = tapply(x[ , varSignal ], x$mac, mean)
                    y = matrix(avgSS, nrow = 1, ncol = 6,  #reshape for our training data, number of columns dependent on macs, this refers to unique number of macs within the mac column
                               dimnames = list(ans$posXY,
                                               names(avgSS)))
                    cbind(ans, y)
                  }))

  newDataSS = do.call("rbind", byLocation)
  return(newDataSS)
}


```

## Create summary level detail, which takes the mean of every position, angle, and mac combination

```{r}

# Create new combined position variable
offline$posXY = paste(offline$posX, offline$posY, sep="-")
byLocAngleAP = with(offline,
                  by (offline, list (posXY, angle, mac),
                      function(x) x))

# Create a summary of signal strength data - takes combination each unique combination of posxy. angle, mac address to create scalar values of means, medians, etc
# Create a summary of signal strength data
signalSummary = 
  lapply(byLocAngleAP, 
        function(oneLoc) 
        {
          ans = oneLoc[1, ]
          ans$medSignal = median(oneLoc$signal)
          ans$avgSignal = mean(oneLoc$signal)
          ans$num = length(oneLoc$signal)
          ans$sdSignal = sd(oneLoc$signal)
          ans$iqrSignal = IQR(oneLoc$signal)
          ans
        })

# Add signalSummary to the offlineSummary data
offlineSummary = do.call("rbind", signalSummary)
head(offlineSummary)
```

## Test Data

## Clean up our test set, this means rounding orientations, removing ,:= , summarizing data, and changing from long to short
* clean our data again with readData
* online - cleaned version
* onlinesummary - summary level, which translates to the mean of every signal for every mac

```{r}
#Cleans up our data w/ readData
online = readData(filename = "Data/online.final.trace.txt", submacs  = co)

angle_count <- online %>% 
  count(angle)

print(angle_count)

print(str(online))

online$posXY = paste(online$posX, online$posY, sep = "-")

keepVars = c("posXY", "posX","posY", "orientation", "angle")
byLoc = with(online, 
             by(online, list(posXY), 
                function(x) {
                  ans = x[1, keepVars]
                  avgSS = tapply(x$signal, x$mac, mean)
                  y = matrix(avgSS, nrow = 1, ncol = 6, #this is co_cd in submacs
                        dimnames = list(ans$posXY, names(avgSS)))
                  cbind(ans, y)
                }))

onlineSummary = do.call("rbind", byLoc)  
```

## selectTrain function

* function that Subsets our training data to include only specific number of angles and also applies reshapeSS which changes from long to short

```{r}

#Create function that will give us starting angle and range (m)
selectTrain = function(angleNewObs, signals , m){
  # m is the number of angles to keep between 1 and 5
  refs = seq(0, by = 45, length  = 8)
  nearestAngle = roundOrientation(angleNewObs)
  
  if (m %% 2 == 1) 
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
  else {
    m = m + 1
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
    if (sign(angleNewObs - nearestAngle) > -1) 
      angles = angles[ -1 ]
    else 
      angles = angles[ -m ]
  }
  angles = angles + nearestAngle
  angles[angles < 0] = angles[ angles < 0 ] + 360
  angles[angles > 360] = angles[ angles > 360 ] - 360
  angles = sort(angles) 
  
  offlineSubset = signals[ signals$angle %in% angles, ]
  reshapeSS(offlineSubset, varSignal = "avgSignal")
}

```

## KNN function-----------------
* trains on our subset summary level data, 

```{r}
findNN = function(newSignal, trainSubset) {
  diffs = apply(trainSubset[ , 4:9], 1, #looks at the signal data, 10 got cd & co
                function(x) x - newSignal)
  dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) ) #takes differences of signal and squares
  closest = order(dists) #reorders differences
  return(trainSubset[closest, 1:3 ]) #keeps top 3 
}

```

## Prediction Function

```{r}
predXY = function(newSignals, newAngles, trainData, 
                  numAngles = 1, k = 3){
  
  closeXY = list(length = nrow(newSignals))
  
  for (i in 1:nrow(newSignals)) {
    trainSS = selectTrain(newAngles[i], trainData, m = numAngles)
    closeXY[[i]] = 
      findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
  }

  estXY = lapply(closeXY, 
                 function(x) sapply(x[ , 2:3], 
                                    function(x) mean(x[1:k])))
  estXY = do.call("rbind", estXY)
  return(estXY)
}

```

```{r actual knn models}
estXYk1 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 1)

estXYk3 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 3)

estXYk5 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 5)

estXYk7 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 7)
```

## KNN Test 2 Results: Removing cd Mac ID

```{r pred error}
# Error calculations
calcError = 
  function(estXY, actualXY) 
    sum( rowSums( (estXY - actualXY)^2) )

actualXY = onlineSummary[ , c("posX", "posY")]
sapply(list(estXYk7, estXYk5, estXYk3, estXYk1), calcError, actualXY)
```

## KNN Test 3: Removing c0 Mac ID

```{r}
offline = readData(filename = "Data/offline.final.trace.txt", 
                   submacs  = cd)
```

## reshapeSS Function
* function to reshape our data from long to short, creates column of average signal per mac
* note, we apply this function after we take summary level details

```{r}
reshapeSS = function(data, varSignal = "signal", 
                     keepVars = c("posXY", "posX","posY")) {
  byLocation =
    with(data, by(data, list(posXY), 
                  function(x) {
                    ans = x[1, keepVars]
                    avgSS = tapply(x[ , varSignal ], x$mac, mean)
                    y = matrix(avgSS, nrow = 1, ncol = 6,  #reshape for our training data, number of columns dependent on macs, this refers to unique number of macs within the mac column
                               dimnames = list(ans$posXY,
                                               names(avgSS)))
                    cbind(ans, y)
                  }))

  newDataSS = do.call("rbind", byLocation)
  return(newDataSS)
}


```

## Create summary level detail, which takes the mean of every position, angle, and mac combination

```{r}

# Create new combined position variable
offline$posXY = paste(offline$posX, offline$posY, sep="-")
byLocAngleAP = with(offline,
                  by (offline, list (posXY, angle, mac),
                      function(x) x))

# Create a summary of signal strength data - takes combination each unique combination of posxy. angle, mac address to create scalar values of means, medians, etc
# Create a summary of signal strength data
signalSummary = 
  lapply(byLocAngleAP, 
        function(oneLoc) 
        {
          ans = oneLoc[1, ]
          ans$medSignal = median(oneLoc$signal)
          ans$avgSignal = mean(oneLoc$signal)
          ans$num = length(oneLoc$signal)
          ans$sdSignal = sd(oneLoc$signal)
          ans$iqrSignal = IQR(oneLoc$signal)
          ans
        })

# Add signalSummary to the offlineSummary data
offlineSummary = do.call("rbind", signalSummary)
head(offlineSummary)
```

## Test Data

## Clean up our test set, this means rounding orientations, removing ,:= , summarizing data, and changing from long to short
* clean our data again with readData
* online - cleaned version
* onlinesummary - summary level, which translates to the mean of every signal for every mac

```{r}
#Cleans up our data w/ readData
online = readData(filename = "Data/online.final.trace.txt", submacs  = cd)

angle_count <- online %>% 
  count(angle)

print(angle_count)

print(str(online))

online$posXY = paste(online$posX, online$posY, sep = "-")

keepVars = c("posXY", "posX","posY", "orientation", "angle")
byLoc = with(online, 
             by(online, list(posXY), 
                function(x) {
                  ans = x[1, keepVars]
                  avgSS = tapply(x$signal, x$mac, mean)
                  y = matrix(avgSS, nrow = 1, ncol = 6, #this is co_cd in submacs
                        dimnames = list(ans$posXY, names(avgSS)))
                  cbind(ans, y)
                }))

onlineSummary = do.call("rbind", byLoc)  
```

## selectTrain function

* function that Subsets our training data to include only specific number of angles and also applies reshapeSS which changes from long to short

```{r}

#Create function that will give us starting angle and range (m)
selectTrain = function(angleNewObs, signals , m){
  # m is the number of angles to keep between 1 and 5
  refs = seq(0, by = 45, length  = 8)
  nearestAngle = roundOrientation(angleNewObs)
  
  if (m %% 2 == 1) 
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
  else {
    m = m + 1
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
    if (sign(angleNewObs - nearestAngle) > -1) 
      angles = angles[ -1 ]
    else 
      angles = angles[ -m ]
  }
  angles = angles + nearestAngle
  angles[angles < 0] = angles[ angles < 0 ] + 360
  angles[angles > 360] = angles[ angles > 360 ] - 360
  angles = sort(angles) 
  
  offlineSubset = signals[ signals$angle %in% angles, ]
  reshapeSS(offlineSubset, varSignal = "avgSignal")
}

```

## KNN function-----------------
* trains on our subset summary level data, 

```{r}
findNN = function(newSignal, trainSubset) {
  diffs = apply(trainSubset[ , 4:9], 1, #looks at the signal data, 10 got cd & co
                function(x) x - newSignal)
  dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) ) #takes differences of signal and squares
  closest = order(dists) #reorders differences
  return(trainSubset[closest, 1:3 ]) #keeps top 3 
}

```

## Prediction Function

```{r}
predXY = function(newSignals, newAngles, trainData, 
                  numAngles = 1, k = 3){
  
  closeXY = list(length = nrow(newSignals))
  
  for (i in 1:nrow(newSignals)) {
    trainSS = selectTrain(newAngles[i], trainData, m = numAngles)
    closeXY[[i]] = 
      findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
  }

  estXY = lapply(closeXY, 
                 function(x) sapply(x[ , 2:3], 
                                    function(x) mean(x[1:k])))
  estXY = do.call("rbind", estXY)
  return(estXY)
}

```

```{r actual knn models}
estXYk1 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 1)

estXYk3 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 3)

estXYk5 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 5)

estXYk7 = predXY(newSignals = onlineSummary[ , 6:11], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 7)
```

## KNN Test 3 Results: Removing c0 MacID

```{r pred error}
# Error calculations
calcError = 
  function(estXY, actualXY) 
    sum( rowSums( (estXY - actualXY)^2) )

actualXY = onlineSummary[ , c("posX", "posY")]
sapply(list(estXYk7, estXYk5, estXYk3, estXYk1), calcError, actualXY)
```

### Clustering Model 2: Weighted Average 

```{r} 
findNN = function(newSignal, trainSubset) {
  diffs = apply(trainSubset[ , 4:9], 1, 
                function(x) x - newSignal) 
  dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) ) 
  closest = order(dists) 
  closeXY = trainSubset[closest, 1:3 ]
  weight = as.numeric(1/dists[closest]) 
  return(cbind(closeXY, weight)) 
}
```

# Weighted Avg. Prediction

```{r} 
predXY = function(newSignals, newAngles, trainData, 
                  numAngles = 1, k = 3){
  
  closeXY = list(length = nrow(newSignals))
  
  for (i in 1:nrow(newSignals)) {
    trainSS = selectTrain(newAngles[i], trainData, m = numAngles)
    original = 
      findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
    wts = append(original[1:k, 4]/sum(original[1:k, 4]), rep(0, nrow(original)-k))
    original[ , 2:3] = original[ , 2:3]*wts 
    closeXY[[i]] = original[ ,1:3] 
  }

  estXY = lapply(closeXY, 
                 function(x) sapply(x[ , 2:3], 
                                    function(x) sum(x)))
  estXY = do.call("rbind", estXY)
  return(estXY)
}

estXYk5w = predXY(newSignals = onlineSummary[ , 6:12], 
                 newAngles = onlineSummary[ , 4], 
                 offlineSummary, numAngles = 3, k = 5)
``` 

# Weighted Average Results

```{r} 

# Error calculations
calcError = 
function(estXY, actualXY) 
   sum( rowSums( (estXY[, 1:2] - actualXY)^2) )

actualXY = onlineSummary[ , c("posX", "posY")]
sapply(list(estXYk5w, estXYk5), calcError, actualXY)
```


