---
title: "unit10_boston-housing"
author: "Martin Garcia, Michael Catalano, Jeremy Otsap, Christian Nava"
date: "6/26/2020"
output: 
  html_document:
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(VIM)
library(skimr)

```

## Boston Data Set 

Using the Boston Housing Data Set examine the effect on loss for data Missing Completely at Random for 10% 20% and 50% of the data. Repeat the analysis for Missing at random and Missing Not at random. 


MCAR
MAR
Monotone


**https://www.kaggle.com/puxama/bostoncsv** 
https://www.kaggle.com/c/boston-housing
https://www.cs.toronto.edu/~delve/data/boston/bostonDetail.html


##Initial Data Exploration

**Data Description**

The Boston data frame has 506 rows and 14 columns.

This data frame contains the following columns:
  
* CRIM - per capita crime rate by town
* ZN - proportion of residential land zoned for lots over 25,000 sq.ft.
* INDUS - proportion of non-retail business acres per town.
* CHAS - Charles River dummy variable (1 if tract bounds river; 0 otherwise)
* NOX - nitric oxides concentration (parts per 10 million)
* RM - average number of rooms per dwelling
* AGE - proportion of owner-occupied units built prior to 1940
* DIS - weighted distances to five Boston employment centres
* RAD - index of accessibility to radial highways
* TAX - full-value property-tax rate per $10,000
* PTRATIO - pupil-teacher ratio by town
* B - 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town
* LSTAT - % lower status of the population
* MEDV - Median value of owner-occupied homes in $1000's


**Importing Data**

```{r, echo=FALSE}

# loading data set

boston.df <- read.csv("https://raw.githubusercontent.com/jotsap/DS7333/master/data/boston.csv", header = T)

```



**Cursory Data Exploration**

```{r, echo=FALSE}

#explore dataframe

str(boston.df)
head(boston.df)
summary(boston.df)


```


Skim allows a quick visual examination of each variable's range

```{r, echo=FALSE}

#skim dataframe
skim(boston.df)


```



**Missing Data**

No missing values found

```{r, echo=FALSE}


# from VIM package
aggr(boston.df, 
     prop = FALSE, 
     combined = TRUE, 
     numbers = TRUE, 
     sortVars = TRUE, 
     sortCombs = TRUE)


```



##Initial Regression Model

We are doing a regression on full data set, predicting MEDV, the median value of owner-occupied homes, using the other 13 parameters.


**Exclude vs Omit**

Using na.exclude pads the residuals and fitted values with NAs where there were missing values. Other functions do not use the na.action, but instead have a different argument (with some default) for how they will handle missing values. For example, the mean command will, by default, return NA if there are any NAs in the passed object.


```{r, echo=FALSE}

# initial full regression

boston.lm <- lm( MEDV ~ . , data = boston.df, na.action = na.omit  )

summary(boston.lm)

```

Improvements for different imputation methods can be measured with various statistics. Adjusted R-square or coefficient of determination can help us assess the models overall quality. This value would help us explain the proportion of variance of median house prices. Fitting a linear regression model without imputation results in a R-square of .7338. The accuracy of the model can be captured by looking at our residuals or actual median house prices minus predicted house prices. The baseline model results in a total sum of square errors of 11078. The goal will be to have a large R-square and small sum of squares error value.
 
```{r}
plot(boston.lm$fitted, boston.lm$residuals, xlab = "Fitted Values", ylab = "Residuals")
abline(0,0)

deviance(boston.lm)

```

**RESULTS**

For "goodness of fit" we ;ppl at the adjusted R-squared value [which penalizes for large numbers of predictors, and get a score of 0.7338. Which means roughly 73.4% of the MEDV response variable is accounted for by the other 13 predictors.


NOTE: Because AGE and INDUS were found to not be statistically significant we will *not* use those for the missing data exercises



Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)  3.646e+01  5.103e+00   7.144 3.28e-12 ***
CRIM        -1.080e-01  3.286e-02  -3.287 0.001087 ** 
ZN           4.642e-02  1.373e-02   3.382 0.000778 ***
INDUS        2.056e-02  6.150e-02   0.334 0.738288    
CHAS         2.687e+00  8.616e-01   3.118 0.001925 ** 
NOX         -1.777e+01  3.820e+00  -4.651 4.25e-06 ***
RM           3.810e+00  4.179e-01   9.116  < 2e-16 ***
AGE          6.922e-04  1.321e-02   0.052 0.958229    
DIS         -1.476e+00  1.995e-01  -7.398 6.01e-13 ***
RAD          3.060e-01  6.635e-02   4.613 5.07e-06 ***
TAX         -1.233e-02  3.760e-03  -3.280 0.001112 ** 
PTRATIO     -9.527e-01  1.308e-01  -7.283 1.31e-12 ***
B            9.312e-03  2.686e-03   3.467 0.000573 ***
LSTAT       -5.248e-01  5.072e-02 -10.347  < 2e-16 ***





##MCAR


First generate vector of *unique* random numbers to correspond to each percentage

* 1% - 5 entries
* 5% - 25 entries
* 10% - 51 entries
* 20% - 102 entries
* 33% - 168 entries
* 50% - 253 entries 


**Value to be Deleted**
* PTRATIO - pupil-teacher ratio by town



Create vector to serve as sample index for creating NA values


```{r, echo=FALSE}

# replace set to FALSE ensures no repeating numbers in sample vector

# 1% - 5 entries
base::sample(1:506, 5, replace = F) -> index_01

# 5% - 25 entires
base::sample(1:506, 25, replace = F) -> index_05

# 10% - 50 entires
base::sample(1:506, 51, replace = F) -> index_10

# 20% - 100 entries
base::sample(1:506, 102, replace = F) -> index_20

# 33% - 165 entries
base::sample(1:506, 168, replace = F) -> index_33

# 50% - 250 entries 
base::sample(1:506, 253, replace = F) -> index_50

```



**Create different boston samples with NA for PTRATIO**


1% - 5 entries 

```{r, echo=FALSE}

# create separate file
boston.df -> boston_mcar_01.df
# assign NA to PTRATIO
boston_mcar_01.df[index_01,'PTRATIO'] <- NA
# validate results

head(boston_mcar_01.df[index_01,'PTRATIO'])

aggr(boston_mcar_01.df, 
     prop = FALSE, 
     combined = TRUE, 
     numbers = TRUE, 
     sortVars = TRUE, 
     sortCombs = TRUE)

```


5% - 25 entries

```{r, echo=FALSE}

# create separate file
boston.df -> boston_mcar_05.df
# assign NA to PTRATIO
boston_mcar_05.df[index_05,'PTRATIO'] <- NA
# validate results

head(boston_mcar_05.df[index_05,'PTRATIO'])

aggr(boston_mcar_05.df, 
     prop = FALSE, 
     combined = TRUE, 
     numbers = TRUE, 
     sortVars = TRUE, 
     sortCombs = TRUE)

```


10% - 51 entries

```{r, echo=FALSE}

# create separate file
boston.df -> boston_mcar_10.df
# assign NA to PTRATIO
boston_mcar_10.df[index_10,'PTRATIO'] <- NA
# validate results

head(boston_mcar_10.df[index_10,'PTRATIO'])

aggr(boston_mcar_10.df, 
     prop = FALSE, 
     combined = TRUE, 
     numbers = TRUE, 
     sortVars = TRUE, 
     sortCombs = TRUE)

```


20% - 102 entries

```{r, echo=FALSE}

# create separate file
boston.df -> boston_mcar_20.df
# assign NA to PTRATIO
boston_mcar_20.df[index_20,'PTRATIO'] <- NA
# validate results

head(boston_mcar_20.df[index_20,'PTRATIO'])

aggr(boston_mcar_20.df, 
     prop = FALSE, 
     combined = TRUE, 
     numbers = TRUE, 
     sortVars = TRUE, 
     sortCombs = TRUE)

```


33% - 168 entries

```{r, echo=FALSE}

# create separate file
boston.df -> boston_mcar_33.df
# assign NA to PTRATIO
boston_mcar_33.df[index_33,'PTRATIO'] <- NA
# validate results

head(boston_mcar_33.df[index_33,'PTRATIO'])

aggr(boston_mcar_33.df, 
     prop = FALSE, 
     combined = TRUE, 
     numbers = TRUE, 
     sortVars = TRUE, 
     sortCombs = TRUE)

```


50% - 253 entries

```{r, echo=FALSE}

# create separate file
boston.df -> boston_mcar_50.df
# assign NA to PTRATIO
boston_mcar_50.df[index_50,'PTRATIO'] <- NA
# validate results

head(boston_mcar_50.df[index_50,'PTRATIO'])

aggr(boston_mcar_50.df, 
     prop = FALSE, 
     combined = TRUE, 
     numbers = TRUE, 
     sortVars = TRUE, 
     sortCombs = TRUE)

```






```{r, echo=FALSE}

# 


```









##MAR


**Determining Parameter for Deletion**
* 10%: TAX <= 233 – [54 values 10.67%]
* 20%: TAX <= 270 – [99 values 19.565%]
* 30%: TAX <= 287 – [149 values 29.447%]

**Values to be Deleted**
* CRIM - per capita crime rate by town
* RAD - index of accessibility to radial highways


###MAR 10%

```{r, echo=FALSE}

# create 10% MAR dataframe
boston.df -> boston_mar_10.df

# list rows with TAX <= 233 
boston_mar_10.df[boston_mar_10.df$TAX <= 233,c("CRIM","RAD")]



```



```{r, echo=FALSE}


#assign NA to CRIM for 10% MAR
boston_mar_10.df$CRIM[boston_mar_10.df$TAX <= 233] <- NA

#assign NA to RAD for 10% MAR
boston_mar_10.df$RAD[boston_mar_10.df$TAX <= 233] <- NA



```



```{r, echo=FALSE}

# list rows with missing values
boston_mar_10.df[is.na(boston_mar_10.df$CRIM),c("CRIM","RAD","TAX")]

# list rows with TAX <= 233 
boston_mar_10.df[boston_mar_10.df$TAX <= 233, c("CRIM","RAD")]

```










###MAR 20%



```{r, echo=FALSE}

# create 10% MAR dataframe
boston.df -> boston_mar_20.df

# list rows with TAX <= 233 
boston_mar_20.df[boston_mar_20.df$TAX <= 270,c("CRIM","RAD")]



```



```{r, echo=FALSE}


#assign NA to CRIM for 10% MAR
boston_mar_20.df$CRIM[boston_mar_20.df$TAX <= 270] <- NA

#assign NA to RAD for 10% MAR
boston_mar_20.df$RAD[boston_mar_20.df$TAX <= 270] <- NA



```



```{r, echo=FALSE}

# list rows with missing values
boston_mar_20.df[is.na(boston_mar_20.df$CRIM),c("CRIM","RAD","TAX")]

# list rows with TAX <= 233 
boston_mar_20.df[boston_mar_20.df$TAX <= 270, c("CRIM","RAD")]

```









###MAR 30%



```{r, echo=FALSE}

# create 10% MAR dataframe
boston.df -> boston_mar_30.df

# list rows with TAX <= 233 
boston_mar_30.df[boston_mar_30.df$TAX <= 287,c("CRIM","RAD")]



```



```{r, echo=FALSE}


#assign NA to CRIM for 10% MAR
boston_mar_30.df$CRIM[boston_mar_30.df$TAX <= 287] <- NA

#assign NA to RAD for 10% MAR
boston_mar_30.df$RAD[boston_mar_30.df$TAX <= 287] <- NA



```



```{r, echo=FALSE}

# list rows with missing values
boston_mar_30.df[is.na(boston_mar_30.df$CRIM),c("CRIM","RAD","TAX")]

# list rows with TAX <= 233 
boston_mar_30.df[boston_mar_30.df$TAX <= 287, c("CRIM","RAD")]

```



















```{r, echo=FALSE}

# 


```






```{r, echo=FALSE}

# 


```



