---
title: "unit10_boston-housing"
author: "Martin Garcia, Michael Catalano, Jeremy Otsap, Christian Nava"
date: "6/26/2020"
output: 
  html_document:
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(VIM)

```

## Boston Data Set 

Using the Boston Housing Data Set examine the effect on loss for data Missing Completely at Random for 10% 20% and 50% of the data. Repeat the analysis for Missing at random and Missing Not at random. 


MCAR
MAR
Monotone


**https://www.kaggle.com/puxama/bostoncsv** 
https://www.kaggle.com/c/boston-housing
https://www.cs.toronto.edu/~delve/data/boston/bostonDetail.html


## Data description

The Boston data frame has 506 rows and 14 columns.

This data frame contains the following columns:
  
* CRIM - per capita crime rate by town
* ZN - proportion of residential land zoned for lots over 25,000 sq.ft.
* INDUS - proportion of non-retail business acres per town.
* CHAS - Charles River dummy variable (1 if tract bounds river; 0 otherwise)
* NOX - nitric oxides concentration (parts per 10 million)
* RM - average number of rooms per dwelling
* AGE - proportion of owner-occupied units built prior to 1940
* DIS - weighted distances to five Boston employment centres
* RAD - index of accessibility to radial highways
* TAX - full-value property-tax rate per $10,000
* PTRATIO - pupil-teacher ratio by town
* B - 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town
* LSTAT - % lower status of the population
* MEDV - Median value of owner-occupied homes in $1000's




```{r, echo=FALSE}

# loading data set

boston.df <- read.csv("https://raw.githubusercontent.com/jotsap/DS7333/master/data/boston.csv", header = T)

```



```{r, echo=FALSE}

#explore dataframe

str(boston.df)
head(boston.df)
summary(boston.df)


```



## Finding Missing Data

No missing values found

```{r, echo=FALSE}


# from VIM package
aggr(boston.df, 
     prop = FALSE, 
     combined = TRUE, 
     numbers = TRUE, 
     sortVars = TRUE, 
     sortCombs = TRUE)


```




